---
title: "HW4 Group 1, Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Oct 1 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---

\newpage

```{r message=FALSE, warning=FALSE, include=FALSE}
# setup
library(tidyverse)
library(stargazer)
library(wooldridge)
```
# Problems

## Question 1

### (i)

I am not surprised by the sign, because the further away from the incinerator I am, the more value I get from my house. As you add a percent of distance away from the incinerator, your home value $log\hat(price)$ would increase by .132%

### (ii)

Other factors that could influence your home prices could be square footage, age of the home, location (good neighborhood, commercial versus residential proximity, etc), or the lot size of the home. These factors could also easily be correlated with distance.

### (iii)

No, because our unobserved variables are not independent of our the price of our home. The city decides as to where the incinerator is built, which would mean that it was placed away from more expensive homes, this would violate our 4th SLR assumption and give us biased results.

## Question 2

### (i)


$$E(u|inc) = E(\sqrt{inc} * e|inc) = (\sqrt{inc})E(e|inc) = (\sqrt{inc})E(e) = (\sqrt{inc}) * 0 = 0$$

### (ii)

$$Var(u|inc)=Var(\sqrt{inc} * e|inc = (\sigma_e)^2*inc$$

### (iii)

As family income increases, the basic needs of the family are more fully met and that family has more discretion as to where they invest their money. Low income families must spend their money on necessities like food, rent, and bills. High income families may choose to invest in savings or spend it on non-necessities, which could explain the variance of increase income to savings.

---

# Computer Problems

## Question 3

```{r echo=TRUE, comment=NA}
## Load the data
rich <- k401k
```


### (i)
```{r echo=TRUE, comment=NA}
mean(rich$prate)
mean(rich$mrate)
```

### (ii)
```{r echo=TRUE, comment=NA}
reg <- lm(prate~mrate, rich)
stargazer(reg, type="text")
```

This gives us an estimated equation of:

$$\hat{prate}=83.075+5.86mrate$$
Our sample size would be: $n=1534$ and $R^2=0.75$


### (iii)

Interpret the intercept and the coefficient of *mrate*

Our intercept means that even if our *mrate* was 0, we would still have a participation rate of 83.05%. Additionally, it means that for every one dollar increase in the match rate, we would see a 5.86% increase in *prate*. 

### (iv)

```{r echo=TRUE, comment=NA}
predict_rich <- function(rate){
  83.075+5.86*rate
}

predict_rich(3.5)
```

When *mrate = 3.5* we get a predicted *prate* of 103.59. This can't happen, because you can't have a participation rate higher than 100%. This is an example that sometimes the simple regression can return strange predictions for with extreme values.

### (v)

Looking at our coefficient estimate, *mrate* explains about 7.5% of the variation in *prate*. This is a low explanation of our model and suggests that there are probably other factors that influence 401(k) plan participation in companies.

\newpage

## Question 4

```{r echo=TRUE, comment=NA}
## Load the data
math <- meap93
```

### (i)

I would say that a diminishing return would make more sense. As you increase spending, the effect of each dollar will not have a constant effect on the pass rate. At one point you may have every student earning passing marks, so an additional dollar would have no effect on a student passing or not.

### (ii)



### (iii)

```{r echo=TRUE, comment=NA}
math_reg_log <- lm(math10~lexpend, math)
stargazer(math_reg_log, type="text")
```
This gives us an estimated equation of:

$$\hat{math10}=-69.34+11.16\ log(expend)$$

Our sample size would be: $n=408$ and $R^2=0.030$

### (iv)

If *expend* increases by 10%, then $\hat{math10}$ increases by about 1.1%. This is not a significant effect, but for schools who spend on average lower amounts, a 10% increase may be a small amount of money.

### (v)

```{r echo=TRUE, comment=NA}
pander::pander(summary(math$math10))
pander::pander(summary(predict(math_reg_log)))
```

If we look at the values for *math10* we see that the maximum value in our data set is 66.70. Furthermore, if we use our model using the fitted values, we still only return a max value of 30.15, so we shouldn't worry about getting values greater than 100 in this dataset.

## Question 5

### (i)

```{r echo=TRUE, comment=NA}
giftsdb <- charity

mean(giftsdb$gift)
# The average gift amount was 7.44 guilders

pander::pander(prop.table(table(giftsdb$respond)))
#From our sample, about 60% of people did not give a gift.
```

### (ii)

```{r echo=TRUE, comment=NA}
# The average mailings per year was
mean(giftsdb$mailsyear)

min(giftsdb$mailsyear)
max(giftsdb$mailsyear)
```
The average mailings per year was `2.05` \newline
The min was `0.25` and the max was `3.5`

### (iii)

```{r echo=TRUE, comment=NA}
gift_reg <- lm(gift~mailsyear, giftsdb)
stargazer(gift_reg, type="text")
```

Our estimated equation is:

$$\hat{gift}=2.014\ +\ 2.65mailsyear$$
We have a $n=4268$ and an $R^2=0.014$

### (iv)

We can interpret our slope to mean that, on average, each additional mailing is associated with 2.65 additional guilders. So if each mailing costs one guilder, they can expect to make 1.65 guilders on each mailing. However, considering that this is only on average. There are many instances where mailings generated no gifts, or where the gift was less than 1 guilder.  

## Question 6

### (i)
```{r echo=TRUE, comment=NA}
## Load the data
cat_stats <- catholic

nrow(cat_stats)
## Number of observations is 7430

mean(cat_stats$math12)
mean(cat_stats$read12)
## The average for Math Scores was 51.13. For Reading 51.77

sd(cat_stats$math12)
sd(cat_stats$read12)
# The standard deviation for Math Scores was 9.46. For Reading 9.41
```

### (ii)

```{r echo=TRUE, comment=NA}
cat_reg <- lm(math12~read12, cat_stats)
stargazer(cat_reg, type="text")
```

Our estimated equation would be:

$$\hat{math12} = 15.153 + 0.714read12$$
Our n is $n=7430$
And $R^2=0.505$

### (iii)

We can interpret our intercept to mean that if our reading score was zero, than you would have a math score of 15.153. This is not a meaningful interpretation since we would not expect to find zero reading score within our observed scores of students.

### (iv)

The $\hat{\beta_1}$ we found was 0.714. This is what I expect we would find, since we would assume that if a student did well in one subject, they are likely to do well in the other and vice versa. With that in mind, we can also expect our $R^2$ value to be high as well based off this assumption. We can see that there is a correlation relationship, but not a causal, as we are not taking into account other important factors in our analysis.

### (v)

Considering that we are only looking at math and reading scores, we are leaving out a lot of other factors that could influence the scores of math. If we switch the analysis and test the effects of math scores on reading, we get a very similar model. There are unobserved factors in our analysis that are not being accounted for which is leading to a bias in our study.
 