---
title: "HW2_Group #1_Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Sep 15 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---



\newpage

# Problems

## Question 1

### (i) 

```{r echo=FALSE, comment=NA}
v1 <- seq(1,17,4)
v1
```

## Question 2

### (i)

```{r echo=FALSE, comment = NA}
my_data <- 20:39
header_names <- LETTERS[1:5]
A <- matrix(my_data, nrow = 4, ncol = 5, byrow = TRUE) 
A
```

### (ii)

```{r echo=FALSE, comment = NA}
colnames(A) <- header_names
A
```

### (iii)

```{r echo=FALSE, comment=NA}
# Extracting Columns B & D and rows 2 & 4
B <- A[c(2,4),c("B","D")]
B
```

### (iv)

```{r echo=FALSE, comment=NA}
t(B)
```

### (v)

```{r echo=FALSE, comment=NA}
# Inverse of Matrix B

B_inverse <- solve(B)
B_inverse
```

### (vi)

```{r echo=FALSE, comment=NA}
B %*% B_inverse
```
 
\newpage 
 
## Question 3

### (i)

```{r echo=FALSE, comment=NA}
df <- data.frame(mtcars)
head(df,10)
```

### (ii)

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
library(dplyr)
df1 <- df %>% 
  select(c(cyl, hp, wt, vs, am, gear, mpg)) %>% 
  arrange(cyl) 
head(df1,5)
```

### (iii)

```{r echo=FALSE, comment=NA}
df1 <- df1 %>% 
  mutate(gpm = round((1/df1$mpg),3)) 
head(df1,5)
```

### (iv)

```{r echo=FALSE, comment=NA}
str(df1)
```

### (v)

```{r echo=FALSE, comment=NA}
names <- c('vs', 'am', 'cyl', 'gear')
df1[,names] <- lapply(df1[,names] , factor)
str(df1)
```

---------



# Computer Exercises

```{r message=FALSE, warning=FALSE, comment=NA, include=FALSE}
library(wooldridge)
df2 <- wooldridge::alcohol
df2 <- df2 %>% 
  select(abuse, status) 
```

## Alcohol Dataset for Questions

### (i)
**What is percentage of the men in the sample report abusing alcohol?**
```{r echo=FALSE, comment=NA}
prop.table(table(df2$abuse))
```
From the sample, we see about 9.9% of men report alcohol abuse

### (ii)
**What is the employment rate?**
```{r echo=FALSE, comment=NA}
employ <- df2 %>% 
  filter(status == c(2,3))

prop.table(table(employ$status))
table(employ$status)
```
The employment rate is ≈ 96.7%

### (iii)
**Consider the group of men who abuse alcohol. What is the employment rate?**
```{r echo=FALSE, comment=NA}
ab_emp <- df2 %>% 
  filter(abuse==1) %>% 
  filter(status != 1)

prop.table(table(ab_emp$status))
```
The employment rate of those who abuse alcohol is ≈ 95.4%

### (iv)
**What is the employment rate for the group of men who do not abuse alcohol?**
```{r echo=FALSE, comment=NA}
non_ab <- df2 %>% 
  filter(abuse==0) %>%
  filter(status!=1)

prop.table(table(non_ab$status))
```
The employment rate for those who do not abuse alcohol is ≈ 96.7%

### (v)

__Discuss the difference in your answers to parts (iii) and (iv). Does this allow you to conclude that alcohol abuse causes unemployment?__

---
```{r echo=FALSE, comment=NA}

```

The rates of those who abuse alcohol and those who do not were who were employed were x% and y%. This does not highlight anything to do with the causality of alcohol abuse and unemployment. We would need to test and control other explanatory variables to try to explain the reasoning for unemployement.
