---
title: "HW3_Group1_Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Sep 22 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---



\newpage

# Problems

## Question 1

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
library(dplyr)
library(tidyverse)
library(wooldridge)

## Make the data frame
family <- data.frame(family = 1:10, monthly_expense = c(300,440,350,1100,640,480,450,700,670,530))
```


### (i)

Find the average monthly housing expenditure.
```{r echo=FALSE, comment=NA}
mean(family$monthly_expense)
```

### (ii)

Find the median monthly housing expenditure.
```{r echo=FALSE, comment=NA}
median(family$monthly_expense)
```

### (iii)

If monthly housing expenditures were measured in hundreds of dollars, rather than in dollars, what would be the average and median expenditures?
```{r echo=FALSE, comment=NA}
adj1 <- family %>% 
  mutate(
    hundreds_dollar = monthly_expense / 100
  ) 

mean(adj1$hundreds_dollar) 
median(adj1$hundreds_dollar)
```
The mean would be 5.66

The median would be 5.05

### (iv)

Suppose that family number 8 increases its monthly housing expenditure to $900, but the expenditures of all other families remain the same. Compute the average and median housing expenditures.
```{r echo=FALSE, message=TRUE, comment=NA}
adj2 <- family
adj2[8,2] <- adj2[8,2] + 100

# Mean
mean(adj2$monthly_expense)
# Median
median(adj2$monthly_expense)
```
The mean would be 576.

The median would be 505.

\newpage

## Question 2

Reconcile the shareholders and CEO's disagreement:
```{r echo=FALSE, comment=NA}
(18-15)/15

# The percentage change from 15% to 18% represents a 20% increase.
```


The difference between the two percentages (18% - 15%) is 3%, however the percentage change between 15% and 18% is 20%. They are both correct, but looking at the problem from two different positions.

## Question 3 

```{r echo=FALSE, comment=NA}
log_salary <- 10.6 + 0.027*(0)
calc_salary <- function(x){
  exp(10.6 + 0.027 * x)
}
```


### (i)

What is salary when exper = 0? When exper = 5? (Hint: You will need to
exponentiate.)

```{r echo=FALSE, comment=NA}
sal0 <- calc_salary(0)
sal5 <- calc_salary(5)
sal0
sal5
```
The salary at 0 years would be $40,134.84. 

The salary at 5 years would be $45,935.80.

### (ii)
Use this equation: %∆y ≈ (100 ∗ β1)∆x to approximate the percentage increase in salary when exper increases by five years.
```{r echo=FALSE, comment=NA}
calc_percent <- function(x){
  (100*0.027)*x
}

calc_percent(5)
```


### (iii)

Use the results of part (i) to compute the exact percentage difference in salary when exper = 0 and When exper = 5. Comment on how this compares with the approximation in part (ii).
```{r echo=FALSE, comment=NA}
((sal5-sal0)/sal0)*100
```
The  percent change is 14.5% from the part i. Based on our approximation from part ii, we had that after 5 years experience the percentage increase would be about 13.5%. 

\newpage 

## Question 4

Let X be a random variable distributed as X ∼ Norm(5,4). Find the probabilities of the following events. Hint: E(X) = 5 and V ar(X) = 4

### (i)

P(X ≤ 6)

```{r echo=FALSE, comment=NA}
round(pnorm(6,5,2),3)
```

### (ii)

P(X > 4)

```{r echo=FALSE, comment=NA}
round(1-pnorm(4,5,2),3)
```

### (iii)

P(|X−5|>1)

```{r echo=FALSE, comment=NA}
round(1-pnorm(6,5,2) + pnorm(4,5,2),3)
```

## Question 5



```{r echo=FALSE, comment=NA}
# Writing the function
cdf <- function(x){
  3*x^2 - 2*x^3
} 
```

### (i)

Find the probability that the elderly employment rate is at least 0.6 (i.e. 60%).

```{r echo=FALSE, comment=NA}
1-cdf(0.6)
```

## Question 6

### (i)

Find the mean and standard deviation when salary is measured in dollars.

```{r echo=FALSE, comment=NA}
# use expected values property

52.3*1000

14.6*1000
```



