---
title: "HW3-G1-ChaseDeanHarward"
author: "Chase Dean Harward"
date: "9/22/2020"
output: word_document
---
\newpage

```{r}
library(wooldridge)
library(dplyr)
```

# Question 1

```{r}
Q1 <- data.frame(family = 1:10, monthly_house_expend = c(300, 440, 350, 1100, 640, 480, 450, 700, 670, 530))

Q1
```


## (i)

```{r}
mean(Q1$monthly_house_expend)
```

## (ii)

```{r}
median(Q1$monthly_house_expend)
```

## (iii)

```{r}
Q1iii <- mutate(Q1, new_expend = monthly_house_expend / 100)
mean(Q1iii$new_expend)
median(Q1iii$new_expend)
```

## (iv)
```{r}
iv <- Q1 
iv[8,2] <- iv[8,2] + 200
mean(iv$monthly_house_expend)
median(iv$monthly_house_expend)
```


# Question 2
The CEO is correct. The stock return did not increase by 3%. It did increase from 15% to 18%, which is a difference of 3%. The CEO is right because the stock return did increase by 20%. 15% increased by 20% is 18%

# Question 3

## (i)
```{r}
salary <- function(exper){
  
  exp(10.6 + (.027 * exper))
  
}

salary(0)
salary(5)
```

## (ii)
```{r}
percent_salary <- function(exper){
  (0.027 * 100) *exper
}
percent_salary(5)
```
13.5% increase

## (iii)
```{r}
((salary(5)-salary(0))) / salary(0) * 100
```
So there was an exact increase of ~14.454% here, while in part (ii) the approximate increase was 13.5%. This is ~1% difference between the two.

# Question 4

## (i)

```{r}
pnorm(6, 5, 2)
```

## (ii)
```{r}
1-pnorm(4,5,2)
```

## (iii)
```{r}
pnorm(4,5,2)+(1-pnorm(6,5,2))
```

# Question 5
```{r}
CDF <- function(x){
  
  3*x^2 - 2*x^3

}

1-CDF(0.6)
```
35.2% probability

# Question 6
Mean
```{r}
52.3 * 1000
```
Standard Deviation
```{r}
14.6 * 1000
```

