---
title: "HW3_Group1_Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Sep 22 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---



\newpage

# Problems

## Question 1

```{r echo=FALSE, comment=NA}
library(dplyr)
library(tidyverse)
library(wooldridge)

## Make the data frame
family <- data.frame(family = 1:10, monthly_expense = c(300,440,350,1100,640,480,450,700,670,530))
```


### (i)

Find the average monthly housing expenditure.
```{r echo=FALSE, comment=NA}
mean(family$monthly_expense)
```

### (ii)
```{r echo=FALSE, comment=NA}
median(family$monthly_expense)
```

### (iii)
```{r echo=FALSE, comment=NA}
adj1 <- family %>% 
  mutate(
    hundreds_dollar = monthly_expense / 100
  ) 

mean(adj1$hundreds_dollar) 
median(adj1$hundreds_dollar)
```


### (iv)
```{r echo=FALSE, comment=NA}
adj2 <- family
adj2[8,2] <- adj2[8,2] + 100
mean(adj2$monthly_expense)
median(adj2$monthly_expense)
```

\newpage

## Question 2
```{r echo=FALSE, comment=NA}
(18-15)/15*100
```


The difference between the two percentages was 3%, however the percentage change between 15% and 18% is 20%.

## Question 3 

```{r echo=FALSE, comment=NA}
log_salary <- 10.6 + 0.027*(0)
calc_salary <- function(x){
  exp(10.6 + 0.027 * x)
}
```


### (i)
```{r echo=FALSE, comment=NA}
sal0 <- calc_salary(0)
sal5 <- calc_salary(5)
sal0
sal5
```


### (ii)
```{r echo=FALSE, comment=NA}
calc_percent <- function(x){
  (100*0.027)*x
}

calc_percent(5)
```


### (iii)
```{r echo=FALSE, comment=NA}
((sal5-sal0)/sal0)*100
```
The exact percent change is 14.5% from the part i. Based on our approximation from part ii, we had that after 5 years experience the percentage increase would be about 13.5%. 

## Question 4

### (i)
```{r echo=FALSE, comment=NA}
round(pnorm(6,5,2),3)
```

### (ii)
```{r echo=FALSE, comment=NA}
round(1-pnorm(4,5,2),3)
```

### (iii)
```{r echo=FALSE, comment=NA}
round(1-pnorm(6,5,2) + pnorm(4,5,2),3)
```

## Question 5

```{r echo=FALSE, comment=NA}
# Writing the function
cdf <- function(x){
  3*x^2 - 2*x^3
} 

1-cdf(0.6)
```

### (i)
```{r echo=FALSE, comment=NA}
1-cdf(0.6)
1-pnorm(0.6)
```

## Question 6

### (i)
```{r echo=FALSE, comment=NA}
# use expected values property

52.3*1000

14.6*1000



```



