---
title: "Group 1"
author: "Austin Halvorsen, Chase Harward, Suneet Dash"
date: "12/8/2020"
output:
  revealjs::revealjs_presentation:
    theme: blood
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(wooldridge)
library(car)
library(effects)
library(corrgram)
library(wooldridge)
library(lmtest)
library(corrplot)
```

```{r data, include=FALSE}
df <- wooldridge::sleep75
df <- df[complete.cases(df),]
df <- select(df, -c(case, leis1, leis2, leis3, hrwage, lothinc, rlxall, south, union, yngkid))
df <- df %>% mutate(sleep_h = sleep/60)

<<<<<<< HEAD
# Data frame of only the variables we are using in the regression for the slide
model_df <- select(df, c(sleep_h, age, exper, lhrwage, yrsmarr, spsepay, male, gdhlth, totwrk))
model <- lm(sleep_h~age+exper+lhrwage+yrsmarr+spsepay+male+gdhlth+totwrk, df)
=======
# Dataframe of only the variables we are using in the regression for the slide
model_df <- select(df, c(sleep_h, age, exper, hrwage, yrsmarr, spsepay, male, gdhlth, totwrk, marr))
model <- lm(sleep_h~hrwage+exper+age+yrsmarr+spsepay+male+gdhlth+totwrk+marr, df)
>>>>>>> refs/remotes/origin/master
```


# Introduction {data-transition="slide-in fade-out"}

__How does someone's hourly wage impact the amount of sleep that they get on a given night?__

<section style="text-align: left;">

- Determining what role hourly wage plays in sleep is important because it can help us to understand an aspect of sleep that isn’t explored as often as opposed to just looking at hours of work itself. 

- Through this question we can even see how different income classes sleep, such as how many hours does someone in the top income percentile sleep as opposed to someone in the middle class or someone making minimum wage.

</section>

## Preconcieved Beliefs

When we started, we expected...

## Why is this Important? {data-transition="slide-in fade-out"}

- Understanding if wage plays a role in a persons sleep could help us identify groups that may be in true need of additional support through public programs, funding or other resources.

- Understanding that if wage plays a key in role in an individuals sleep could indicate other aspects of quality of life. This could also be indicative of personal preference, i.e. some people may just choose to sleep more or less than others.

# Data {data-transition="slide-in fade-out"}

- <small>We will be using the dataset “sleep75” found in the Wooldridge package and originally used for the paper “Sleep and the Allocation of Time” by Jeff Biddle and Daniel Hamermesh, published in the Journal of Political Economy in 1998.</small>

- <small>The goal of the original paper was to show a relationship between an increase in time in the labor market and how much sleep someone gets.</small>
<<<<<<< HEAD

- <small>This data set includes 706 observations on 34 variables, notably hourly wage, sleep, years of experience, marital status, spouse wage, age, children or no children, type of employment, and quality of health.</small>

## Variables

Our explanatory variables in our model are:

<section style="text-align: left;">

- age            
- exper           
- lhrwage
- yrsmarr         
- spsepay        
- male         
- gdhlth

</section>

## Variables (con't)

Our Dependent

## Summary Stats 

```{r summary, echo=FALSE, comment=NA}
(summary(model_df))
```

# Emperical Framework {data-transition="slide-in fade-out"}

```{r}
summary(wooldridge::sleep75)
```


- <small>This data set includes 706 observations on 34 variables, notably hourly wage, sleep, years of experience, marital status, spouse wage, age, children or no children, type of employment, and quality of health.</small>

## Variables

Our variables in our model are:

<section style="text-align: left;">

- age            
- exper           
- lhrwage
- yrsmarr         
- spsepay        
- male         
- gdhlth

</section>

## Summary Stats 

```{r echo=FALSE, comment=NA}
(summary(model_df))
```

## Classical Linear Model (CLM) Assumptions

1. MLR.1 Linear in Parameters
2. MLR.2 Random Sampling
3. MLR.3 No Perfect Collinearity
4. MLR.4 Zero Conditional Mean
5. MLR.5 Homoskedasticity
6. MLR.6 Normality

## MLR.1 Linear in Parameters

- All of our parameters are linear

## MLR.2 Random Sampling

<<<<<<< HEAD
- Data is taken from Wooldridge dataset sleep75, which was randomly sampled with 706 observations
=======
- Data is taken from wooldridge dataset sleep75, which was randomly sampled.
>>>>>>> refs/remotes/origin/master


## MLR.3 No Perfect Collinearity
```{r echo=FALSE, comment=NA}

```



## MLR.4 Zero Conditional Mean

- We have included the variables from the data set that we intuitively believe to have a correlation with the amount of sleep that someone gets in a week. 
- Refer possibly to Pedram’s comment about him not yet showing us how to numerically prove the zero conditional mean assumption

## MLR.5 Homoskedasticity

```{r echo=FALSE, comment=NA}
#Breusch-Pagan test
bptest(model)
```
- We can prove that the error has the same variance across all values of the explanatory variables 
- According to the results of our Breusch-Pagan test, we have a p value of .1939, and so we fail to reject the null hypothesis

## MLR.6 Normality

```{r echo=FALSE, comment=NA}
hist(resid(model), freq =FALSE)
lines(density(resid(model), adjust = 2), lty = "dotted", col="red", lwd=3) 
```

- One of the most important assumptions for this analysis, and we can prove that there is normality.

# Results {data-transition="slide-in fade-out"}

<small>

- $B^4loghrwage$: on average, a 1% increase in hourly wage is associated with a decrease in hours of sleep per week of 0.002191 (1 second per day or 7.89 seconds per week)
- $R^2$ is .02305 or 2.305%
- $\text{Adjusted } R^2$ is .008103 or .81%

- The variation in these variables explain very little of the variation in our dependent variable sleep
- T statistic for log hrly wage is -.371, it is not statistically significant for a 5% critical value
- Show test of heteroskedasticity
- Even when testing for join significance, p value is only .1398, still insignificant at a 5% significance level or even a 10% significance level

</small>

## Hypothesis Test

<small>

- We are asking whether or not there is a statistically significant relationship between hourly wage and hours of sleep per week

- $H_0$ is that $B^4 = 0$ (hourly wage tells us very little about predicting someone’s sleep)
- $H_a$ is that $B^4 \text{ does not} = 0$ (hourly wage may have some effect on sleep)
- P value is __.7111__-smallest significance level that it is statistically significant at is 71.11%
    - This is a very high p value
- With a rejection region of <5% and a t value of -0.371, we fail to reject the null hypothesis
    - We have a very small t value
- There is a 95% chance that the true population coefficient for B^4 lies within:
    - (-1.338, 1.118) since 0 lies within the confidence interval, we fail to reject the null hypothesis

</small>
