---
title: "HW7 Group 1, Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Nov 3 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---

\newpage

# Problems

```{r message=FALSE, warning=FALSE, include=FALSE}
# setup
library(tidyverse)
library(stargazer)
library(wooldridge)
library(pander)
library(car)
```

## Question 1

### (i) 

If you have heteroskedasticity, you violate assumption MLR5 and cannot rely on the t stat

### (ii)

If you have a sample correlation coefficient of .95 between two independent variables in the model, you are not in violation of the assumptions and you can use the t stat.

### (iii)

If you omit an important explanatory variable, you introduce bias into your model, which would violate MLR.4, and you would not be able to rely on the t test for your model.

\newpage

## Question 2

### (i) 

The null hypothesis would be $H_0:\beta_3=0$. The alternative hypothesis would be $H_1:\beta_3\neq0$

### (ii)

I would expect $\beta_1$ to have a positive sign, since having a larger population would result in more demand for rental housing.

I would expect $\beta_2$ to also have a positive sign, since higher income would likely mean that the cost of living is high as well, meaning housing prices and rental rates would be higher.

### (iii)

Because this is a log-log model, the interpretation of this model should be, *"A 10% increase in population would result an increase of rent of 0.66%.

### (iv)

With a degree of freedom of 60 (64-4), at the 1% level, we need to calculate the t value as $$t=\frac{\hat{\beta_3}-\beta_{3_{H_0}}}{se(\hat{\beta_3})}$$
This would give an estimated t stat of $\frac{0.0056-0}{0.0017}≈3.294$
```{r}
qt(0.995, 60)
```

We would reject the null hypothesis at the 1% significance level.

\newpage

## Question 3

### (i)

We fail to reject the null hypothesis where $\beta_0=0$ and we also fail to reject the null hypothesis where $\beta_1=1$. We calculate this be evaluating the t stat for each parameter: 

$$\beta_0[\,tstat]\,=\left\lvert\frac{-14.47}{16.27}\right\rvert≈0.889$$ and $$\beta_1[\,tstat]\,= \left\lvert \frac{0.976-1}{.049} \right\rvert≈0.4898$$

### (ii)

We need to calculate the F statistic where:

$$F=\frac{\frac{(SSR_R-SSR_{UR})}{q}}{\frac{SSR_{UR}}{n-k-1}} = \frac{\frac{209448.99-165644.51}{2}}{\frac{165644.51}{86}}=11.3713$$
So we are going to reject the null that they are jointly significant.

### (iii)

To test the joint hypothesis that $H_0:\beta_2=0, \beta_3=0, and \space \beta_4=0$ we need to test

$$F=\frac{\frac{(R^2_{UR}-R^2_R)}{q}}{\frac{(1-R^2_{UR})}{n-k-1}} = \frac{\frac{0.829-0.820}{3}}{\frac{1-0.829}{88-4-1}}=1.45614$$
So, with this output, we fail to reject the null.

### (iv) 

You cannot rely on the F statistic from part iii because we have violated assumption MLR.5. 

\newpage

## Question 4

### (i)

We can apply the property of variance: 
$$Var(\hat{\theta})=Var(\hat{\beta}_1-3\hat{\beta}_2)=Var(\hat{\beta}_1)+9Var(\hat{\beta}_2)-6 Cov(\hat{\beta}_1,\hat{\beta}_2)$$

### (ii)

$$t = \frac{(\hat{\beta}_1-3\hat{\beta}_2-1)}{se(\hat{\beta}_1-3\hat{\beta}_2)}$$

So to identify the t statistic, we get:

$$t = \frac{(\hat{\beta}_1-3\hat{\beta}_2-1)}{se[\,(\hat{\beta}_1)+9Var(\hat{\beta}_2)-6 Cov(\hat{\beta}_1,\hat{\beta}_2)]\,}$$

### (iii)

If $\theta_1=\beta_1-3\beta_2$ then $\beta_1=\theta_1+3\beta_2$, therefore we can write our population regression model as:

$$y=\beta_0+(\theta_1+3\beta_2)x_1+\beta_2x_2+\beta_3x_3+u$$
$$=\beta_0+\theta_1x_1+\beta_2(3x_1+x_2)+\beta_3x_3+u$$
We would estimate y by regressing y on $x_1, 3x_1+x_2,$ and $x_3$. This will allow us to calculate the stand error and coefficient of $x_1$

\newpage

## Question 5

### (i)

With our t stat close to 1: ($\frac{.0023}{.0022}$), we will fail to reject that this variable is irrelevant.

### (ii)

Yes, the effect is significant. The t value is greater than 2, and a critical value of and this would make sense that companies with larger market values pay more. 

### (iii)

The t value for ceoten is $\frac{.0171}{.0055}≈3.101$ and the t value for comten $\frac{.0092}{.0033}≈2.788$ so they are both significant to our model.

### (iv)

This could be the result of someone who has stayed with the company for a long time and may not have demands on pay and does the job more out of passion for the company rather than an outsider who may demand large bonus pay to run the company.

---------

\newpage

# Computer Exercises

## Question 6

### (i)
```{r comment=NA}
d1 <- hprice1
mrm1 <- lm(lprice~sqrft+bdrms, d1)
stargazer(mrm1, type = 'text', digits = 8)
```

Using the data from HPRICE1, we get the following model:

$$log(price)=4.766+0.0003795sqrft+0.02889bdrms$$ with an $n=88$ and $R^2=0.588$.

Using our model we get the following estimation: 
$$\hat{\theta_1}=150(.0003795)+0.02889=0.0858$$
```{r comment=NA}
theta <- 150*mrm1$coefficients[2]+mrm1$coefficients[3]
theta
```

This means that when we add a 150 sqft bedroom, then we can expect the home price to increase by about 8.6%

### (ii)

We can rewrite $\beta_2=\theta_1-150\beta_1$. In our model, we can substitute to have the following:

$$log(price)=\beta_0+\beta_1(sqrft-150bdrms)+\theta_1bdrms + u$$

### (iii)

```{r comment=NA}
se1 <- lm(lprice~I(sqrft-150*bdrms)+bdrms, d1)
stargazer(se1, type = 'text')

pander(confint(se1, level = 0.95))
```



## Question 7

### (i)

```{r comment=NA}
d2 <- k401ksubs
d2  %>% 
  filter(fsize==1) %>% 
  count()
```

There are 2017 single person households in the dataset.

### (ii)

```{r comment=NA}
d2a <- d2 %>% 
  filter(fsize==1)
mrm2 <- lm(nettfa~inc+age, d2a)
stargazer(mrm2, type='text')
```

Using only single family households, we get the model:

$$nettfa=-43.040+0.799inc+0.843age$$
With an $n=2017$ and $R^2=0.119$. This means that for every \$1 thousand increase in income means an estimated \$799 increase in net wealth. For $\beta_2$ we interpret that for every year increase, we estimate a \$843 net wealth increase. These numbers make sense, and is not too surprising.

### (iii)

The intercept, $\hat{\beta_0}=-43.040$, means that at age 0 and income = 0, their net wealth is $-43.04. In other words, you are born in debt. 

### (iv)

```{r comment=NA}
pt(-1.5,2014)
```


### (v)

```{r comment=NA}
srm <- lm(nettfa~inc, d2a)
stargazer(srm, mrm2, type = 'text')
```

When calculating the regression using just inc, the estimated coefficient is larger than it is the first model. It is going to be larger because the correlation is small between age and inc. So when we remove the age, we introduce a small postivie bias into the model which gives us a larger estimated coefficient.

## Question 8

### (i)

```{r comment=NA}
d3 <- discrim
mrm3 <- lm(lpsoda~prpblck+lincome+prppov, d3)
stargazer(mrm3, type='text')
```

We get the following model:

$$log(psoda)=-1.463+0.073prpblck+0.137log(income)+0.380prppov$$

With an $n=401$ and $R^2=0.087$. AT the 5% level, $\hat{\beta_1}$ is statistically significant, but it is not significant at the 1% level.

### (ii)

```{r comment=NA}
pander(cor.test(d3$lincome,d3$prppov, method = "pearson", use="complete.obs"))
summary(mrm3)
```

The correlation between log(income) and prppov is **-0.8385**.

The p-value for log(income) is 0.0045 and the p-value for prppov is 0.00001. These are both significant at the 1% level.

### (iii)

```{r comment=NA}
mrm4 <- lm(lpsoda~prpblck+lincome+prppov+lhseval, d3)
stargazer(mrm4, type='text')
```

Adding the variable **log(hseval)** we get the following model:

$$log(psoda)=-0.842+0.098prpblck-0.053log(income)+0.052prppov+0.121log(hseval)$$
Using this model, we can interpret the coefficient to mean that a 1% higher median housing value in a zip code with increase the predicted price of a medium soda by 0.121%.

For the two-sided p-value at $H_0:\beta_{log(hseval)}=0$ we have a low p-value, less than 

