---
title: "HW7 Group 1, Austin Halvorsen"
author: "Pedram Jahangiry"
date: "Nov 1 2020"
output:
  word_document: 
    keep_md: yes
  pdf_document: default
---

\newpage

# Problems

```{r message=FALSE, warning=FALSE, include=FALSE}
# setup
library(tidyverse)
library(stargazer)
library(wooldridge)
library(pander)
library(car)
```

## Question 1

### (i) 

### (ii)

### (iii)

## Question 2

### (i) 

The null hypothesis would be $H_0: \beta_3=0$. The alternative hypothesis would be $H_1: \beta_3\neq0$

### (ii)

I would expect $\beta_1$ to have a positive sign, since having a larger population would result in more demand for rental housing.

I would expect $\beta_2$ to also have a positive sign, since higher income would likely mean that the cost of living is high as well, meaning housing prices and rental rates would be higher.

### (iii)

Because this is a log-log model, the interpretation of this model should be, *"A 10% increase in population would result an increase of rent of 0.66%.

### (iv)

With a degree of freedom of 60 (64-4), at the 1% level, we 

## Question 3

### (i)

### (ii)

### (iii)

### (iv) 

\newpage

## Question 4

### (i)

We can apply the property of variance: 
$$Var(\hat{\beta}_1-3\hat{\beta}_2=Var(\hat{\beta}_1)+9Var(\hat{\beta}_2)-6 Cov(\hat{\beta}_1,\hat{\beta}_2)$$

### (ii)

$$t = \frac{(\hat{\beta}_1-3\hat{\beta}_2-1)}{se(\hat{\beta}_1-3\hat{\beta}_2)}$$

We will need to identify the standard error, or $se(\hat{\beta}_1-3\hat{\beta}_2)$

### (iii)

If $\theta_1=\beta_1-3\beta_2$ then $\beta_1=\theta_1+3\beta_2$, therefore we can write our population regression model as:

$$y=\beta_0+(\theta_1+3\beta_2)x_1+\beta_2x_2+\beta_3x_3+u$$
$$=\beta_0+\theta_1x_1+\beta_2(3x_1+x_2)+\beta_3x_3+u$$
We would estimate y by regressing y on $x_1, 3x_1+x_2,$ and $x_3$. This will allow us to calculate the stand error and coefficient of $x_1$

## Question 5

### (i)

### (ii)

### (iii)

### (iv)

### (v)

---------

\newpage

# Computer Exercises

## Question 6

### (i)
```{r comment=NA}

```

### (ii)

### (iii)

### (iv)

### (v)


## Question 7

### (i)

### (ii)

### (iii)

### (iv)

### (v)


## Question 8

### (i)

### (ii)

### (iii)
